---
title: "Prototype"
---

# Data Preparation 1

A specific tab will be imported to do the prototype: Market Share of Electricity Generation

```{r}
pacman::p_load(scales, viridis, lubridate, ggthemes, gridExtra, readxl, knitr, data.table, CGPfunctions, ggHoriPlot, tidyverse,sf, tmap)
```

```{r}
SES <- read_excel("data/SES_Public_2023_tidy.xlsx", sheet = "T2.6")
```

```{r}
str(SES)
```

```{r}
unique(SES$company)
```

# Sample Visualisation 1: Boxplot

```{r,warning=FALSE}
# Assuming 'year' is a column in your SES dataframe
filtered_data <- subset(SES, year == 2023)

# Drop rows with NA values
filtered_data <- na.omit(filtered_data)

# Define a mapping of current names to new names
filtered_data <- mutate(filtered_data, company = recode(
  company, "Tuas Power Generation" = "Tuas",
                   "Senoko Energy" = "Senoko",
                   "YTL PowerSeraya" = "YTL",
                   "Keppel Merlimau Cogen" = "Keppel",
                   "SembCorp Cogen" = "Sembcorp",
                   "PacificLight Power" = "PacificLight",
                   "Taser Power" = "Taser",
                   "Others" = "Others"),share_percent = as.numeric(share_percent) )



ggplot(data = filtered_data, aes(x = reorder(company, -share_percent), y = share_percent))  +
  geom_bar(stat = "identity") +
  ggtitle("Market Share of Electricity Generation Companies - 2023") +
  xlab("Company") +
  ylab("Share Percent")+
  scale_y_continuous(labels = scales::percent_format(scale = 1))+
   theme(axis.title.y= element_text(angle=0), 
        axis.title.y.right = element_text(angle=0, vjust = 1),
        panel.grid.major = element_line(color= 'grey', size = 0.1),
        panel.background= element_blank(), 
        axis.line= element_line(color= 'grey'),
        plot.caption = element_text(hjust = 0))
```

The bar graph can be further improved.

# Data Preparation 2

```{r}
mpsz <- st_read(dsn = "data/geospatial", 
                layer = "MPSZ-2019")
```

```{r}
mpsz
```

Import back the SES Public 2023 data but with another tab

```{r,warning=FALSE}
SES_2 <- read_excel("data/SES_Public_2023_tidy.xlsx", sheet = "T3.5")
```

It is keep to 2023, and overall with annual data

```{r}
SES_2_1 <- SES_2 %>%
  filter(year == 2023, month == "Annual", dwelling_type == "Overall", Description != "Central Region", Description != "North Region", Description != "North East Region",Description != "Overall", Description != "West Region")
  
```

Convert areas to upper case

```{r, warning=FALSE}
SES_2_1 <- SES_2_1 %>%
  mutate_at(.vars = vars(Description), 
          .funs = funs(toupper))
```

Combine geospatial data and SES data

```{r}
mpsz_SES <- left_join(mpsz, SES_2_1,
                          by = c("PLN_AREA_N" = "Description"))
```

Remove rows with na

```{r}
mpsz_SES_filtered <- mpsz_SES %>%
  mutate(kwh_per_acc = coalesce(kwh_per_acc, 0),dwelling_type = coalesce(dwelling_type, "NA"))%>%
  select(dwelling_type,kwh_per_acc, geometry) 
```

# Sample Visualisation 2: Geospatial

```{r,warning=FALSE}
tmap_mode("plot")
tmap_options(check.and.fix = TRUE)
qtm(mpsz_SES_filtered, 
    fill = "kwh_per_acc")


```

```{r,warning=FALSE}
tm_shape(mpsz_SES_filtered)+
  tm_fill("kwh_per_acc", 
          style = "quantile", 
          palette = "Blues",
          title = "Electricity Consumption in kwh_per_acc") +
  tm_layout(main.title = "Total Household Electricity Consumption by Household in 2023",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from Urban Redevelopment Authorithy (URA)\n and electricity data from Energy Marketing Authority (EMA)", 
             position = c("left", "bottom"))
```
